---
title: "Single Cell Tensorflow and BiocSklearn test"
author: "Nitesh Turaga"
date: "6/1/2018"
output: html_document
---

Use virtualenv which is manually installed, and don't depend on auto
created `.virtualenv/r-tensorflow` by tensorflow package.

This virtualenv uses python 3.6

```{r use_private_virtualenv}
library(reticulate)
use_virtualenv("~/Virtualenvs/p3")
```

Load `tensorflow` and `BiocSklearn` packages as both python libraries
are installed in this private virtualenv. The next chunk works only if
the `use_virtualenv()` function is pointed to the correct path.

```{r load_libraries}
library(tensorflow)
library(BiocSklearn)
library(SingleCellExperiment)
```

Test successful loading of packages by writing test functions.

**Basics of tensorflow is that first we create a model which is called
a computational graph with tensorflow objects then we create a
tensorflow session in which we start running all the computation**.

Speed of computation gains are most when tensorflow is used in GPU
settings.


```{r test_tensorflow}
matrix1 <- tf$constant(matrix(c(3.0, 3.0), nrow = 1, ncol = 2))
matrix2 <- tf$constant(matrix(c(3.0, 3.0), nrow = 2, ncol = 1))
product <- tf$matmul(matrix1, matrix2)

## dimensions
product$shape
## name
product$name
## tensorflow object type
product$graph

## Different way of closing session
with(tf$Session() %as% sess, {
  result = sess$run(product)
  print(result)
})

```

Test BiocSklearn

```{r test_biocsklearn}
SklearnEls()
```


Other features to manage python virtualenv from within R,


```{r echo=TRUE, eval=FALSE}
## Install python packages
virtualenv_install("scikit-learn")

## Create new virtualenv under the root
virtualenv_create(envname)

## virtualenv_root needs to be changed 
## Virtual environments are by default located at ~/.virtualenvs. You can change this behavior by defining the WORKON_HOME environment variable.
Sys.setenv("WORKON_HOME = ~/Virtualenvs")

## See function definition
virtualenv_root

## Function definition from 
virtualenv_root <- function() {
  Sys.getenv("WORKON_HOME", unset = "~/.virtualenvs")
}


## List virtual envs (fixed root "~/.virtualenvs")
virtualenv_list()

## Remove packages installed.
virtualenv_remove("scikit-learn")
```


Load single Cell data from the Tabula-muris is the dataset

```{bash download_data, echo=TRUE, eval=FALSE}
## FACS
wget https://ndownloader.figshare.com/files/10038307
unzip 10038307
wget https://ndownloader.figshare.com/files/10038310
mv 10038310 FACS_metadata.csv
wget https://ndownloader.figshare.com/files/10039267
mv 10039267 FACS_annotations.csv

## 10x brain
wget https://ndownloader.figshare.com/files/10038325
unzip 10038325
wget https://ndownloader.figshare.com/files/10038328
mv 10038328 droplet_metadata.csv
wget https://ndownloader.figshare.com/files/10039264
mv 10039264 droplet_annotation.csv
```

```{r load_data_facs}
dat = read.delim("FACS/Kidney-counts.csv", sep = ",", header=TRUE)

dat[1:5, 1:5]

# Remove rownames
rownames(dat) <- dat[,1]
dat <- dat[,-1]

## Check spike ins
rownames(dat)[grep("^ERCC-", rownames(dat))]

cellIDs <- colnames(dat)
cell_info <- strsplit(cellIDs, "\\.")
Well <- vapply(cell_info, `[`, 1, FUN.VALUE = character(1))
Plate <- vapply(cell_info, `[`, 2, FUN.VALUE = character(1))
Mouse <- vapply(cell_info, `[`, 3, FUN.VALUE = character(1))

## Annotation
ann <- read.table("FACS_annotations.csv", sep = ",", header = TRUE)
ann <- ann[match(cellIDs, ann[,1]),]
celltype <- ann[,3]

## Make single cell object
cell_anns <- data.frame(mouse = Mouse, well = Well, type = celltype)
rownames(cell_anns) <- colnames(dat)
facs_sce <- SingleCellExperiment(assays = list(counts = as.matrix(dat)), colData=cell_anns)
```

Read in 10x data for one tissue type and sample.

```{r sce_object_10x}
cellbarcodes <- read.table("droplet/Kidney-10X_P4_5/barcodes.tsv")
genenames <- read.table("droplet/Kidney-10X_P4_5/genes.tsv")
molecules <- Matrix::readMM("droplet/Kidney-10X_P4_5/matrix.mtx")

head(cellbarcodes)
head(genenames)
head(rownames(molecules))

## Give sparse matrix rownames and colnames
rownames(molecules) <- genenames[,1]
colnames(molecules) <- paste("10X_P4_5", cellbarcodes[,1], sep = "_")
```

Get droplet metadata and compare annotation with counts.

```{r droplet_metadata}
metadata <- read.delim("droplet_metadata.csv", sep = ",", header = TRUE)

head(metadata)

## Get metadata for kidney only

metadata[metadata$channel == "10X_P4_5", ]

mouseID <- "3_8_M"

ann <- read.delim("droplet_annotation.csv", sep=",", header=TRUE)
head(ann)

## Fix formatting
ann[,1] <- paste(ann[,1], "-1", sep="")
ann_subset <- ann[match(colnames(molecules), ann[,1]),]
celltype <- ann_subset[,3]

## build metadata dataframe
cell_anns <- data.frame(mouse = rep(mouseID, times=ncol(molecules)), type=celltype)

rownames(cell_anns) <- colnames(molecules)

```

Make a function and read all 10x batches

```{r read_10x_function}

## Function to read in 10x data from droplet folder
read_10x <-
    function(droplet_folder)
{
    cellbarcodes <- read.table(file.path(droplet_folder, "barcodes.tsv"))
    genenames <- read.table(file.path(droplet_folder, "genes.tsv"))
    molecules <- Matrix::readMM(file.path(droplet_folder, "matrix.mtx"))
    
    head(cellbarcodes)
    head(genenames)
    head(rownames(molecules))

    tissue <- unlist(strsplit(droplet_folder, "-"))[1]
    code <- unlist(strsplit(droplet_folder, "-"))[2]
    
    ## Give sparse matrix rownames and colnames
    rownames(molecules) <- genenames[,1]
    colnames(molecules) <- paste(code , cellbarcodes[,1], sep = "_")

    ## metadata
    metadata <- read.delim("droplet_metadata.csv", sep = ",", header = TRUE)
    
    head(metadata)    
    metadata[metadata$channel == code, ]    
    mouseID <- "3_8_M"

    ## annotation
    ann <- read.delim("droplet_annotation.csv", sep=",", header=TRUE)
    head(ann)
    
    ## Fix formatting
    ann[,1] <- paste(ann[,1], "-1", sep="")
    ann_subset <- ann[match(colnames(molecules), ann[,1]),]
    celltype <- ann_subset[,3]
    
    ## build metadata dataframe
    cell_anns <- data.frame(mouse = rep(mouseID, times=ncol(molecules)), type=celltype)

    rownames(cell_anns) <- colnames(molecules)
    SingleCellExperiment(assays=list(counts = as.matrix(molecules)),
                         colData=cell_anns)
}

```

Building a SingleCellExperiemnt object with all kidney cells.

```{r build_sce_object}
molecules1 <- read_10x(droplet_folder = "droplet/Kidney-10X_P7_5")
molecules2 <- read_10x(droplet_folder = "droplet/Kidney-10X_P4_6")
molecules3 <- read_10x(droplet_folder = "droplet/Kidney-10X_P4_5")

## Check if they were read in
molecules1[1:10, 1:10]
molecules2[1:10, 1:10]

## Check if rownames are correct
identical(rownames(molecules1), rownames(molecules2))
identical(rownames(molecules1), rownames(molecules3))

## Check if there are any repeated cellIDs
sum(colnames(molecules1) %in% colnames(molecules2))
sum(colnames(molecules1) %in% colnames(molecules3))
sum(colnames(molecules2) %in% colnames(molecules3))
```

Since everything seems ok, we should combine them into one object
called **all_sce**. This object is saved as `kidney_droplet.rds`.

```{r combine}
all_sce <- cbind(molecules1, molecules2, molecules3)

saveRDS(all_sce, file="kidney_droplet.rds")
```


NMF with Rpackage  NMF

```{r}
library(NMF)

V = counts(all_sce)

res <- NMF::nmf(V, 3)

table(rowSums(V) == 0)
V <- V[rowSums(V) != 0, ] 
w <- basis(res)
h <- coef(res)

```


NMF with BiocSklearn

```{r}
X = counts(all_sce)

nmf <- SklearnEls()$skd$NMF

model <- nmf(n_components=3L, init="random", random_state=0L)

W = model$fit_transform(X)
H = model$components_
```


NMF using tensorflow,

```{r}
library(tensorflow)

## Matrix to be decomposed
A <- counts(all_sce)

## show dimensions of A
dim(A)

A = tf$constant(counts(all_sce))
shape = A$shape
shape

#latent factors (we get to choose this)
rank = 3 

# Initializing random H and W
temp_H = np.random.randn(rank, shape[1]).astype(np.float32)
temp_H = np.divide(temp_H, temp_H.max())

temp_W = np.random.randn(shape[0], rank).astype(np.float32)
temp_W = np.divide(temp_W, temp_W.max())

H =  tf.Variable(temp_H)
W = tf.Variable(temp_W)
WH = tf.matmul(W, H)
```



